---
title: "An Evaluation of Value-at-Risk Forecasting Methods"
output: rmarkdown::github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi=300, fig.height = 4, fig.path = "plots/README/README-")

source("./R/01_import_and_clean.R")

library(tidyverse)
library(forecast)
library(dynlm)
library(cowplot)
library(GGally)
library(lmtest)
library(FinTS)
library(broom)

```

## Introduction 

Volatility permeates the economy. It influences the pricing of assets and plays a prominent role in financial decision-making. One area of application is Value-at-Risk (VaR) modelling, which seeks to quantify the potential financial losses and the probability of their occurance. Firms and regulators alike use VaR as a risk management tool. Since the groundbreaking work of Engle (1982), many different approaches have been developed to model and forecast volatility. This analysis assesses the performance of a wide set of popular univariate volatiltiy models in terms of daily VaR forecasts for 6 major stock market indices, and considers how dimensions such as sample length and parameterization affect VaR forecasts. 

## Data

This analysis considers daily prices of 6 major stock market indices:

   * S&P 500
   * Dow Jones Industrial Average
   * FTSE 100
   * DAX Performance-Index
   * CAC 40
   * Nikkei 225
   
All data is retrieved from Yahoo using the `quantmod` package for the period January 1, 2019 to September 15, 2019. The level as well as daily log-returns of each index is plotted below.
   
**Stock Market Indices Daily Price: 2000 - 2019**
```{r, price_by_index, echo = FALSE, warnings = FALSE, out.width ='100%'}
# Stock Market Indices Daily Price: 2000 - 2019
indices %>% 
   tk_tbl(rename_index = "date") %>% 
   gather(market_index, price, -date) %>% 
   mutate(market_index = fct_relabel(market_index, ~str_remove(., ".Adjusted"))) %>% 
ggplot(aes(x = date, y = price)) +
   geom_line() +
   facet_wrap(~market_index, scale = "free_y") +
   labs(x = NULL,
        y = "Price (local currency)") +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45),
         strip.background = element_rect(fill = "grey90"))
```

**Stock Market Index Returns: 2000 - 2019**
``` {r, log_return_by_index, echo = FALSE, out.width ='100%'}
# Stock Market Index Returns: 2000 - 2019
nested_returns %>% 
   unnest(cols = data) %>% 
   ungroup() %>% 
   mutate(market_index = fct_relabel(market_index, ~str_remove(., ".Adjusted"))) %>% 
ggplot(aes(x = date, y = return)) +
   geom_line() +
   facet_wrap(~market_index) +
   labs(x = NULL,
        y = "Daily Log Return") +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45),
         strip.background = element_rect(fill = "grey90"))
```

Each log-return series exhibits heteroscedasticity or volatiltiy clustering, the phenomenon first described by Mandelbrot (1963): "large changes tend to be followed by large changes of either sign, and small changes tend to be followed by small changes." Periods of high volatility are clearly visible in late 2008 as well as in 2002.

The ACF and PACF plots of each index's squared returns show serial correlations, indicating daily returns are not independent and signalling the presence of ARCH-effects.

```{r, acf_returns_by_index_build, eval = TRUE, include = FALSE}
# names of ACFs to plot
squared_returns<- returns %>% 
   # calculate square returns
   mutate_if(is.numeric,~ .^2)

acf_squared_returns <- map(names(squared_returns)[-1], ~ggAcf(squared_returns[[.]], lag.max = 50) + 
                                          labs(title = NULL,
                                               y = NULL,
                                               subtitle = paste0(str_remove(., ".Adjusted"), " squared returns")))

# sub grid with ACF plots
acf_grid <- plot_grid(plotlist = acf_squared_returns, ncol = 3)

# create title 
title_acf <- ggdraw() + 
   draw_label("ACF of Stock Market Index Squared Returns",
              fontface = 'bold',
              x = 0,
              hjust = 0,
              size = 12
   ) +
   theme(plot.margin = margin(0, 0, 0, 7))
```

```{r, acf_returns_by_index, echo = FALSE, out.width ='100%'}
# plot ACF of squared returns by index
plot_grid(title_acf, acf_grid, ncol = 1, rel_heights = c(0.05, 1))
```

```{r, pacf_squared_returns_build, eval = TRUE, include = FALSE}
pacf_squared_returns <- map(names(squared_returns)[-1], ~ggPacf(squared_returns[[.]], lag.max = 50) + 
                              labs(title = NULL,
                                   y = NULL,
                                   subtitle = paste0(str_remove(., ".Adjusted"), " squared returns")))

# sub grid with ACF plots
pacf_grid <- plot_grid(plotlist = pacf_squared_returns, ncol = 3)

# create title 
title_pacf <- ggdraw() + 
   draw_label("PACF of Stock Market Index Squared Returns",
              fontface = 'bold',
              x = 0,
              hjust = 0,
              size = 12
   ) +
   theme(plot.margin = margin(0, 0, 0, 7))
```

```{r, pacf_squared_returns, echo = FALSE, out.width ='100%'}
# plot ACF of squared returns by index
plot_grid(title_pacf, pacf_grid, ncol = 1, rel_heights = c(0.05, 1))
```

## Models

This analysis considers ...

**GARCH(p,q) of Bollerslev (1986)**

$$r_t = \mu + \epsilon_t ,\ \ \epsilon_t = \sigma_tz_t$$
$$\sigma^2_t = \omega + \sum_{i=1}^{q} \alpha_i \epsilon^2_{t-i} + \sum^p_{j=1} \beta_j \sigma^2_{t-j}$$
where $\simga^2_t$ is the conditional variance, $\omega$ a constant term, and $\epsilon^2_t$ are residuals obtained from the mean filtering process. The ARCH and GARCH orders are determined by $q$ and $p$ respectively. $p = 0$ gives the ARCH(q) model as specified by Engle(1982), while $q > 0$ leads the more general GARCH(p,q) specified by Bollerslev (1986). The analysis here considers only the GARCH(1,1). 

$$z_t \sim iid \ N(0,1)$$

**Threshold GARCH (TGARCH) of Zakoian (1994)**

$$r_t = \mu + \epsilon_t ,\ \ \epsilon_t = \sigma_tz_t$$
$$\sigma_t=\omega + \sum^q_{i=1}\alpha_i(\lvert\epsilon_{t-i}\rvert-\gamma_i\epsilon_{t-i}) + \sum^p_{j=1}\beta_j\sigma_{t-j}  $$
with $\sigma_t$ denoting the conditional standard deviaion. $\omega$ represents a constant term and $\epsilon_t$ are residuals obtained from the mean model. $q$ and $p$ give the ARCH and GARCH orders respectively. 

$$z_t \sim iid \ N(0,1)$$





   
   
   
   
   
   
   
   
   
   
   